### 2024-05-27 out/01. Introduction.tex

\input{out/01. Introduction/00. Preface.tex}
\section{Moving sofa problem}
\label{sec:moving-sofa-problem}
\input{out/01. Introduction/01. Moving sofa problem.tex}

\section{Moving hallway problem}
\label{sec:moving-hallway-problem}
\input{out/01. Introduction/02. Moving hallway problem.tex}

\section{Main result}
\label{sec:main-result}
\input{out/01. Introduction/03. Main result.tex}

\section{Overview of chapters}
\label{sec:overview-of-chapters}
\input{out/01. Introduction/04. Overview of chapters.tex}