### 2024-05-27 out/01. Introduction/02. Moving hallway problem.tex

The moving sofa problem can be seen as an \emph{inverse Kakeya problem} \autocite{bae2021reverse,cabelloInverseKakeyaProblem2022}. The original Kakeya problem asks for the minimum-area chamber \(R\) in which a needle \(S\) of length 1 can be rotated the full \(360\degree\), allowing the needle to translate in its movement \autocite{kakeya1917some}. In an \emph{inverse} Kakeya problem, the shape of the chamber \(R\) is fixed and we are asked to find the shape \(S\) (the ‘needle’) of largest area that can rotate freely inside the chamber, also allowing translations in movement. For example, if the chamber is the strip \(R = \mathbb{R} \times [0, 1]\) of unit width, and the shape \(S\) has to rotate the full \(360\degree\), the problem becomes the well-studied \emph{isodiametric problem} asking for the maximum-area planar shape with diameter at most 1.

The moving sofa problem, as an inverse Kakeya problem, has the hallway as the chamber \(R = L\) and the moving sofa \(S\) as the ‘needle’ inside \(L\). The inverse Kakeya problem has been solved If the chamber \(R\) is a planar convex set; it is a consequence of the Brunn-Minkowski theorem that the largest-area shape \(S\) that can rotate the full \(360\degree\) inside a convex chamber \(R\) is the largest ball inscribed in \(R\) \autocite{cabelloInverseKakeyaProblem2022}. Since the hallway \(L\) is non-convex, however, the same approach is not effective on the moving sofa problem.

The \emph{rotation angle} \(\omega\) of a moving sofa \(S\) is defined as the clockwise angle that \(S\) rotates as it moves from \(L_H\) to \(L_V\) inside \(L\). As an inverse Kakeya problem, the angle \(\omega\) dictates how much the moving sofa \(S\) should rotate inside the hallway \(L\). Gerver’s sofa has the rotation angle \(\omega = \pi/2\). On the other hand, the unit square \([0, 1]^2\) can move inside \(L\) with only translation, so it has the rotation angle \(\omega = 0\).\footnote{As an inverse Kakeya problem, it may occur that smaller value of total rotation angle \(\omega\) will lead to less movement of \(S\) and thus a larger area of \(S\). However, the extra constraints that \(S\) has to move from horizontal side \(L_H\) to vertical side \(L_V\) penalizes the area of \(S\) as \(\omega\) gets smaller from \(\pi/2\).} Let \(\omega\) be the rotation angle of a maximum-area moving sofa. Gerver showed that we can assume \(\pi/ 3 \leq \omega \leq \pi/2\) (\autocite{gerverMovingSofaCorner1992}). Kallus and Romik improved the lower bound of \(\omega\) by showing that \(\omega \geq \arcsin(84/85) = 81.203\dots^\circ\) \autocite{kallusImprovedUpperBounds2018}. With this, we will only consider moving sofas with rotation angle \(\omega \in (0, \pi/2]\), and it seems reasonable to conjecture that a maximum-area moving sofa has \(\omega = \pi/2\).

\begin{conjecture}

(Angle Hypothesis) There exists a maximum-area moving sofa with a movement of rotation angle \(\omega = \pi/2\).

\label{con:angle}
\end{conjecture}

\begin{remark}

A single moving sofa \(S\) may admit different movements with varying rotation angles \(\omega\). For any moving sofa \(S\) mentioned in this paper, we will always assume a fixed rotation angle \(\omega\) attached to it. So any moving sofa in this paper is technically a tuple of a shape and its fixed rotation angle. In this way, we can talk about \emph{the} rotation angle of a moving sofa.

\label{rem:angle}
\end{remark}

Once we fix the rotation angle \(\omega \in (0, \pi/2]\) of a moving sofa \(S\), we can change the moving sofa problem as the moving \emph{hallway} problem. In \autocite{gerverMovingSofaCorner1992}, Gerver looked at the hallway \(L\) in the perspective of the moving sofa \(S\), so that \(S\) is fixed and the hallway \(L\) moves around the sofa (see \Cref{fig:monotone-sofa}). In this perspective, we claim that the sofa \(S\) is a common subset of the following sets.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  For every angle \(t \in [0, \omega]\), the rotated hallway \(L_t\), which is the hallway \(L\) rotated counterclockwise by \(t\) and translated so that the outer walls of \(L_t\) are in contact with \(S\).
\item
  The horizontal strip \(H = \mathbb{R} \times [0, 1]\).
\item
  A translation of \(V_\omega\), where \(V_\omega\) is the vertical strip \(V = [0, 1] \times \mathbb{R}\) rotated counterclockwise by \(\omega\) around the origin.
\end{enumerate}

To observe (1) that \(S \subseteq L_t\) for every angle \(t \in [0, \omega]\), note that the sofa \(S\) is rotated initially by a clockwise angle of \(0\) and finally by \(\omega\) during its movement. By the intermediate value theorem, there is a moment where a copy of \(S\) is rotated clockwise by \(t\) inside \(L\). Push the rotated copy of \(S\) in \(L\) towards the positive \(x\) and \(y\) directions, until it makes contact with the outer walls \(x=1\) and \(y=1\) of \(L\). See this configuration in the perspective of \(S\) to conclude \(S \subseteq L_t\). To observe (2) and (3), note that the sofa \(S\) is initialy in \(L_H\) and finally in \(L_H\) rotated clockwise by \(\omega\); see the configurations in perspective of \(S\) ot conclude that \(S\) is in \(H\) and a translation of \(V_\omega\).

\begin{figure}
\centering
\includesvg[width=1\textwidth,height=\textheight]{images/monotone-sofa-combined.svg}
\caption{The movement of a moving sofa \(S\) with rotation angle \(\omega = \pi/2\) in perspective of the hallway (left) and the sofa (right). The monotone sofa \(S\) is equal to the cap \(K\) minus its niche \(N\). The cap \(K\) is a convex body with the outer walls of \(L_t = L_S(t)\) as the tangent lines of cap \(K\) (and sofa \(S\)). The niche \(N\) is the union of all triangular regions bounded from above by the inner walls of \(L_t = L_S(t)\).}
\label{fig:monotone-sofa}
\end{figure}

Without loss of generality, we can translate the moving sofa \(S\) horizontally and assume that \(S\) is contained in \(V_\omega\), not its arbitrary translation.\footnote{Translating the moving sofa \(S\) may invalidate the initial condition \(S \subseteq L_H\). To resolve this, we modify the \Cref{def:sofa} of a moving sofa \(S\) in so that only some \emph{translation} of \(S\) is required to be movable from \(L_H\) to \(L_V\) inside \(L\).} Then from (1), (2) and (3) above, we can always assume that \(S\) is contained in the intersection
\begin{equation}
\label{eqn:monotone}
\mathcal{M}(S) := H \cap V_\omega \cap \bigcap_{0 \leq t \leq \omega} L_t
\end{equation}
that we will define as the \emph{monotonization} of a moving sofa \(S\). The set \(\mathcal{M}(S)\) is also movable inside \(L\); observe that \(\mathcal{M}(S)\) is in each rotating hallway \(L_t\) of angle \(t \in [0, \omega]\), and see this in the perspective of the fixed hallway \(L\) instead of the moving hallway \(L_t\) and vary \(t\) from \(0\) to \(\omega\). So the monotonization \(\mathcal{M}(S)\) is a moving sofa containing \(S\)\footnote{This uses an implicit assumption that the intersection \(\mathcal{M}(S)\) should be connected. Indeed, \autocite{gerverMovingSofaCorner1992} assumes the connectedness of \(\mathcal{M}(S)\) in its proof of Theorem 1 of \autocite{gerverMovingSofaCorner1992} without any justification. We justify this assumption rigorously in \Cref{thm:monotonization-is-connected}.} (\Cref{thm:monotonization-is-sofa}), and we call such a sofa \(\mathcal{M}(S)\) from monotonization a \emph{monotone sofa} (\Cref{def:monotone-sofa}). Since any moving sofa \(S\) is contained in a larger monotone sofa \(\mathcal{M}(S)\), we only need to consider monotone sofas for the moving sofa problem.