### 2024-05-27 out/15. Monotone sofas.tex

\input{out/15. Monotone sofas/00. Preface.tex}
\section{Tangent Hallway}
\label{sec:tangent-hallway}
\input{out/15. Monotone sofas/07. Tangent Hallway.tex}

\section{Monotonization}
\label{sec:monotonization}
\input{out/15. Monotone sofas/10. Monotonization.tex}

\section{Structure of a monotone sofa}
\label{sec:structure-of-a-monotone-sofa}
\input{out/15. Monotone sofas/12. Structure of a monotone sofa.tex}

\section{Cap contains niche}
\label{sec:cap-contains-niche}
\input{out/15. Monotone sofas/15. Cap contains niche.tex}