### 2024-05-27 out/A1. Convex bodies/00. Preface.tex

Fix an arbitrary planar convex body \(K\), which by \Cref{def:convex-body} is a nonempty, compact and convex subset of \(\mathbb{R}^2\). This appendix defines and proves numerous properties of \(K\). For the ease of understanding, it is recommended to read only the parts of this appendix when needed. However, an interested reader can read this appendix from beginning to end to verify the correctness of all proofs and theorems. Note that we allow \(K\) to have empty interior. If a theorem requires \(K\) to have nonempty interior, we state it explicitly.