### 2024-05-27 front_materials/acknowledgments.tex
fixes

I would like to deeply thank my advisor, Prof. Michael Zieve,
for his support and guidance throughout the program.
His valuable insights, during both hard and good times,
has been instrumental in shaping me into a professional mathematician,
I am also profoundly grateful to Prof. Dan Romik for his strong interest, commitment, and guidance in my research project.
His genuine encouragement was a crucial motivation for me to complete the program,
and his thoughtful feedback significantly improved the presentation of this work.

I would also like to extend my sincere thanks to the rest of my doctoral committee for their time and interest in my work.
In particular, Prof. David Speyer's purely voluntary dedication to understanding the intricacies of my research and 
his help in refining its delivery has been invaluable.
His brilliance, passion, and unconditional support will have a lasting impression on me.
I am also grateful to Prof. Jeffrey Lagarias for his advice during the early stages of my research. 
Additionally, I would like to thank Prof. Joseph Gerver and Prof. Thomas Hales for their appreciation of this project.
Their warm hospitality and voluntary assistance in discussing my work were greatly appreciated.

I am grateful to my friends at Michigan for the memories in Ann Arbor,
that made this town feel like another hometown in the U.S.
To my family, thank you for your unwavering love and support throughout my life;
your presence has been a constant source of motivation.
Finally, to Eseul Choi, your companionship and unconditional love have made life truly worth pursuing.
Thank you for being by my side.