### 2024-05-27 out/25. Conditional upper bound A1.tex
title change

\input{out/25. Conditional upper bound A1/00. Preface.tex}
\section{Definition of $\mathcal{A}_1$}
\label{sec:definition-of-a1}
\input{out/25. Conditional upper bound A1/01. Definition of A1.tex}

\section{Calculus of variation}
\label{sec:calculus-of-variation}
\input{out/25. Conditional upper bound A1/02. Calculus of variation.tex}

\section{Boundary measure}
\label{sec:boundary-measure}
\input{out/25. Conditional upper bound A1/03. Boundary measure.tex}

\section{Concavity of $\mathcal{A}_1$}
\label{sec:concavity-of-a1}
\input{out/25. Conditional upper bound A1/10. Concavity of A1.tex}

\section{Directional derivative of $\mathcal{A}_1$}
\label{sec:directional-derivative-of-a1}
\input{out/25. Conditional upper bound A1/15. Directional derivative of A1.tex}

\section{Maximizer of $\mathcal{A}_1$}
\label{sec:maximizer-of-a1}
\input{out/25. Conditional upper bound A1/20. Maximizer of A1.tex}