### 2024-05-29 out/10. Notations and conventions.tex

\input{out/10. Notations and conventions/00. Preface.tex}
\section{Notations}
\label{sec:notations}
\input{out/10. Notations and conventions/01. Notations.tex}

\section{Moving Sofa}
\label{sec:moving-sofa}
\input{out/10. Notations and conventions/02. Moving sofa.tex}

\section{Convex Body}
\label{sec:convex-body}
\input{out/10. Notations and conventions/03. Convex body.tex}

\section{Area Functionals}
\label{sec:area-functionals}
\input{out/10. Notations and conventions/04. Area functionals.tex}

\section{Parts of Hallway}
\label{sec:parts-of-hallway}
\input{out/10. Notations and conventions/05. Parts of hallway.tex}