### 2024-05-29 front_materials/acknowledgments.tex
fix grammar

I would like to show my deepest gratitude to my advisor, Prof. Michael Zieve,
for his support and guidance throughout the program.
His valuable insights, during both hard and good times,
have been instrumental in my development as a professional mathematician.

I am also profoundly grateful to Prof. Dan Romik for his kindness, dedication, 
and guidance in my research project.
His genuine encouragement was a crucial motivation for me to complete the program,
and his thoughtful feedback with attention to detail
significantly improved the presentation of this work.

I would also like to sincerely thank the rest of my doctoral committee for their time and interest in my work.
In particular, Prof. David Speyer's purely voluntary dedication to understanding the intricacies of my research and 
his help in refining its delivery has been invaluable.
His brilliance, passion, and unconditional support have left a lasting impression on me.

In addition to my committee members,
I am also grateful to Prof. Jeffrey Lagarias for his advice during the early stage of my research
and encouraging remarks throughout the program.
I would like to also thank Prof. Joseph Gerver and Prof. Thomas Hales for their appreciation of this project.
Their warm hospitality and intellectual discussions were greatly appreciated.

I am thankful to my friends in Michigan for the memories in Ann Arbor,
that made this town feel like another hometown in the U.S.
I am also grateful to my friends in Korea for their support.
In particular, I am indebted to Hyunwook Nam for his significant help that
made this research possible.

To my family, thank you for your love and support throughout my life;
your presence has been a constant source of motivation.
Finally, to Eseul Choi, 
your companionship and unconditional love have made life worth pursuing.
Thank you for being by my side.