### 2024-06-15 out/A1. Convex bodies.tex

\input{out/A1. Convex bodies/00. Preface.tex}
\subsection{Vertex and support function}
\label{sec:vertex-and-support-function}
\input{out/A1. Convex bodies/05. Vertex and support function.tex}

\subsection{Lebesgue-Stieltjes measure}
\label{sec:lebesgue-stieltjes-measure}
\input{out/A1. Convex bodies/07. Lebesgue-Stieltjes measure.tex}

\subsection{Surface area measure}
\label{sec:surface-area-measure}
\input{out/A1. Convex bodies/10. Surface area measure.tex}

\subsection{Parametrization of boundary}
\label{sec:parametrization-of-boundary}
\input{out/A1. Convex bodies/15. Parametrization of boundary.tex}

\subsection{Normal angles}
\label{sec:normal-angles}
\input{out/A1. Convex bodies/20. Normal angles.tex}

\subsection{Mamikon's theorem}
\label{sec:mamikon's-theorem}
\input{out/A1. Convex bodies/25. Mamikon's theorem.tex}