### 2024-06-15 out/10._Notations_and_conventions.tex

\input{out/10._Notations_and_conventions/00._Preface.tex}
\subsection{Notations}
\label{sec:notations}
\input{out/10._Notations_and_conventions/01._Notations.tex}

\subsection{Moving Sofa}
\label{sec:moving-sofa}
\input{out/10._Notations_and_conventions/02._Moving_sofa.tex}

\subsection{Convex Body}
\label{sec:convex-body}
\input{out/10._Notations_and_conventions/03._Convex_body.tex}

\subsection{Area Functionals}
\label{sec:area-functionals}
\input{out/10._Notations_and_conventions/04._Area_functionals.tex}

\subsection{Parts of Hallway}
\label{sec:parts-of-hallway}
\input{out/10._Notations_and_conventions/05._Parts_of_hallway.tex}