### 2024-06-15 out/15._Monotone_sofas.tex

\input{out/15._Monotone_sofas/00._Preface.tex}
\subsection{Tangent Hallway}
\label{sec:tangent-hallway}
\input{out/15._Monotone_sofas/07._Tangent_Hallway.tex}

\subsection{Monotonization}
\label{sec:monotonization}
\input{out/15._Monotone_sofas/10._Monotonization.tex}

\subsection{Structure of a Monotone Sofa}
\label{sec:structure-of-a-monotone-sofa}
\input{out/15._Monotone_sofas/12._Structure_of_a_monotone_sofa.tex}

\subsection{Cap Contains Niche}
\label{sec:cap-contains-niche}
\input{out/15._Monotone_sofas/15._Cap_contains_niche.tex}