### 2024-06-15 out/20._Sofa_area_functional_A/00._Preface.tex

We use the findings in \Cref{sec:monotone-sofas} to reduce the moving sofa problem to the maximization of \emph{sofa area functional} \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\) defined on the \emph{space of caps} \(\mathcal{K}_\omega\) with rotation angle \(\omega\). We first define the domain of \(\mathcal{A}\).

\begin{definition}

Define the \emph{space of caps} \(\mathcal{K}_\omega\) with the \emph{rotation angle} \(\omega \in (0, \pi/2]\) as the collection of all caps \(K\) with rotation angle \(\omega\) as in \Cref{def:cap}.

\label{def:cap-space}
\end{definition}

Now define the sofa area functional \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\).

\begin{definition}

For any angle \(\omega \in (0, \pi/2]\), define the \emph{sofa area functional} \(\mathcal{A}_\omega : \mathcal{K}_\omega \to \mathbb{R}\) on the space of caps \(\mathcal{K}_\omega\) as \(\mathcal{A}_\omega(K) = |K| - |\mathcal{N}(K)|\).

\label{def:sofa-area-functional}
\end{definition}

We can restate \Cref{thm:niche-in-cap} using \Cref{def:sofa-area-functional} as:

\begin{corollary}

If \(K \in \mathcal{K}_\omega\) is the cap \(\mathcal{C}(S)\) for a monotone sofa \(S\) with rotation angle \(\omega\), we have \(\mathcal{A}(K) = |S|\).

\label{cor:sofa-area-functional}
\end{corollary}

As we observed in \Cref{rem:niche-not-in-cap}, not all cap \(K \in \mathcal{K}_\omega\) is the cap \(\mathcal{C}(S)\) of a monotone sofa \(S\).

\begin{definition}

For any angle \(\omega \in (0, \pi/2]\), define \(\mathcal{M}_\omega\) as the subset of \(\mathcal{K}_\omega\) of the caps \(\mathcal{C}(S)\) (\Cref{def:cap-sofa}) of an arbitrary monotone sofa \(S\) with rotation angle \(\omega\).

\label{def:monotone-sofa-embedding}
\end{definition}

\(\mathcal{M}_\omega\) is a proper subset of \(\mathcal{K}_\omega\) by \Cref{rem:niche-not-in-cap}. By \Cref{thm:monotone-sofa-structure}, the set of all monotone sofas \(S\) with rotation angle \(\omega\) embeds to the subset \(\mathcal{M}_\omega\) of \(\mathcal{K}_\omega\) by taking the cap \(S \mapsto \mathcal{C}(S)\). By \Cref{thm:monotonization-is-sofa} and \Cref{cor:sofa-area-functional}, the moving sofa problem for a fixed rotation angle \(\omega \in (0, \pi/2]\) is now equivalent to the maximization of the sofa area functional \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\) on the subspace \(\mathcal{M}_\omega\) of \(\mathcal{K}_\omega\).

We will, however, try to optimize the sofa area functional \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\) over the whole \(\mathcal{K}_\omega\), not the subspace \(\mathcal{M}_\omega\) of \(\mathcal{K}_\omega\). This is because the space \(\mathcal{K}_\omega\) of all caps is turns out to be a convex space (\Cref{def:convex-space}) which is easier to understand than the subspace \(\mathcal{M}_\omega\). By extending the domain of optimization of \(\mathcal{A}\) from \(\mathcal{M}_\omega\) to \(\mathcal{K}_\omega\), we get a counterpart of every statement on maximum-area monotone sofas. To start, we have the following strenghtening of Gerver’s conjecture \(\mu_{\max} = \mu_G\).

\begin{conjecture}

The cap \(K = K_G\) of Gerver’s sofa \(S_G\) attains the maximum value \(\mathcal{A}(K)\) of the sofa area functional \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\) over all rotation angle \(\omega \in (0, \pi/2]\).

\label{con:gerver-cap}
\end{conjecture}

While we cannot prove \Cref{con:gerver-cap}, we expect it to be true. The rest of the paper proves the following strenghtening of main \Cref{thm:main} that extends the domain from \(\mathcal{M}_\omega\) to \(\mathcal{K}_\omega\).

\begin{theorem}

(Generalized main theorem) Let \(K \in \mathcal{K}_\omega\) be any cap with rotation angle \(\omega \in (0, \pi/2]\). If the rotation path \(\mathbf{x}_K : [0, \omega] \to \mathbb{R}^2\) of \(K\) is injective and always on the fan \(F_\omega\), then we have \(\mathcal{A}(K) \leq 1 + \omega^2/2\).

\label{thm:main-cap}
\end{theorem}

\begin{proposition}

\Cref{thm:main-cap} implies \Cref{thm:main}.

\label{pro:main-cap}
\end{proposition}

\begin{proof}
Let \(S\) be any monotone sofa of rotation angle \(\omega \in (0 ,\pi/2]\) with cap \(K = \mathcal{C}(S)\), so that \(S\) satisfies the injectivity condition (\Cref{def:injectivity}). By \Cref{pro:cap-same-tangent-hallway} we have \(\mathbf{x}_S = \mathbf{x}_K\). So the injectivity condition of \(S\) implies that \(\mathbf{x}_K : [0, \omega] \to \mathbb{R}^2\) is injective and always on the fan \(F_\omega\). By \Cref{thm:main-cap} we have \(\mathcal{A}(K) \leq 1 + \omega^2/2\). By \Cref{cor:sofa-area-functional} we have \(|S| = \mathcal{A}(K)\), completing the proof.
\end{proof}

We finish this section by mentioning the counterparts of angle and injectivity hypotheses (\Cref{con:angle} and ) that extend the domain from \(\mathcal{M}_\omega\) to \(\mathcal{K}_\omega\).

\begin{conjecture}

The supremum of \(\mathcal{A}_{\omega} : \mathcal{K}_\omega \to \mathbb{R}\) maximizes at \(\omega = \pi/2\).

\label{con:angle-cap}
\end{conjecture}

\begin{conjecture}

There exists a maximizer \(K \in \mathcal{K}_\omega\) of the sofa area functional \(\mathcal{A}_{\omega}\) over all rotation angle \(\omega \in (0, \pi/2]\), such that the rotation path \(\mathbf{x}_K : [0, \omega] \to \mathbb{R}^2\) is injective and always on the fan \(F_\omega\).

\label{con:injectivity-cap}
\end{conjecture}

Perhaps surprisingly, \Cref{con:angle-cap} does not necessarily imply \Cref{con:angle}. Similarly, \Cref{con:injectivity-cap} does not necessarily imply \Cref{con:injectivity}. This is because a maximizer of \(\mathcal{A}_\omega\) is not necessarily a cap of a monotone sofa in \(\mathcal{M}_\omega\). However, observe that proving \Cref{con:angle-cap} would allow us to assume \(\omega = \pi/2\) in a potential proof of \Cref{con:injectivity-cap}, and \Cref{thm:main-cap} with a proof of \Cref{con:injectivity-cap} would imply the upper bound \(1 + \pi^2/8\) of sofa area unconditionally. Proving \Cref{con:angle-cap} and \Cref{con:injectivity-cap} will be the main goal of subsequent works.