### 2024-06-15 out/25._Conditional_upper_bound_A1.tex

\input{out/25._Conditional_upper_bound_A1/00._Preface.tex}
\subsection{Definition of $\mathcal{A}_1$}
\label{sec:definition-of-a1}
\input{out/25._Conditional_upper_bound_A1/01._Definition_of_A1.tex}

\subsection{Calculus of variation}
\label{sec:calculus-of-variation}
\input{out/25._Conditional_upper_bound_A1/02._Calculus_of_variation.tex}

\subsection{Boundary measure}
\label{sec:boundary-measure}
\input{out/25._Conditional_upper_bound_A1/03._Boundary_measure.tex}

\subsection{Concavity of $\mathcal{A}_1$}
\label{sec:concavity-of-a1}
\input{out/25._Conditional_upper_bound_A1/10._Concavity_of_A1.tex}

\subsection{Directional derivative of $\mathcal{A}_1$}
\label{sec:directional-derivative-of-a1}
\input{out/25._Conditional_upper_bound_A1/15._Directional_derivative_of_A1.tex}

\subsection{Maximizer of $\mathcal{A}_1$}
\label{sec:maximizer-of-a1}
\input{out/25._Conditional_upper_bound_A1/20._Maximizer_of_A1.tex}