### 2024-06-15 out/A1._Convex_bodies.tex

\input{out/A1._Convex_bodies/00._Preface.tex}
\subsection{Vertex and support function}
\label{sec:vertex-and-support-function}
\input{out/A1._Convex_bodies/05._Vertex_and_support_function.tex}

\subsection{Lebesgue-Stieltjes measure}
\label{sec:lebesgue-stieltjes-measure}
\input{out/A1._Convex_bodies/07._Lebesgue-Stieltjes_measure.tex}

\subsection{Surface area measure}
\label{sec:surface-area-measure}
\input{out/A1._Convex_bodies/10._Surface_area_measure.tex}

\subsection{Parametrization of boundary}
\label{sec:parametrization-of-boundary}
\input{out/A1._Convex_bodies/15._Parametrization_of_boundary.tex}

\subsection{Normal angles}
\label{sec:normal-angles}
\input{out/A1._Convex_bodies/20._Normal_angles.tex}

\subsection{Mamikon's theorem}
\label{sec:mamikon's-theorem}
\input{out/A1._Convex_bodies/25._Mamikons_theorem.tex}